Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-25 PAPER.tex
updated jackknife explanation

The highest sensitivity power spectrum result for PAPER-64 using the updated analysis presented in this paper, shown in 
Figure \ref{fig:ps1_data}, has positive biases at low $k$ values. As discussed in Section \ref{sec:BiasTypes}, these detections 
are most likely attributable to foreground leakage. Here we demonstrate three null tests performed on PAPER-64 data that verify 
that the positive detections are indeed due to foreground variation and not attributable to EoR.

@@

\begin{itemize}
\item Original Data (black): This is identical to the unweighted, revised PAPER-64 power spectrum in Figure \ref{fig:ps1_data} (one baseline type only) with weighting matrix $\textbf{I}$. There are clear detections at 
low $k$'s.
\item Even/Odd (blue): We split our dataset into even (\textbf{e}) and odd (\textbf{o}) days. We then form the following two datasets: $\textbf{x}_{1} = \textbf{e} + \textbf{o}$ and $\textbf{x}_{2} = 
\textbf{e} - \textbf{o}$. We form datasets in this way to ensure that we use the full sensitivity of our data. When cross-multiplied, we obtain:
\begin{equation}
\textbf{x}_{1}\textbf{x}_{2}^{\dagger} = \textbf{ee}^{\dagger} - \textbf{eo}^{\dagger} + \textbf{oe}^{\dagger} - \textbf{oo}^{\dagger}
\end{equation}
If the same sky signal is in both the even and odd datasets, we expect it to cancel out.
\item Baselines (green): We split our dataset into two halves, where each contains half of the total baselines used in the 
analysis. Again, we form $\textbf{x}_{1} = \textbf{b}_{1} + \textbf{b}_{2}$ and $\textbf{x}_{2} = \textbf{b}_{1} - \textbf{b}_{2}$, where $\textbf{b}_{1}$ is the first baseline 
group and $\textbf{b}_{2}$ is the second baseline group.
\item LST (magenta): We split our dataset into two halves based on LST, namely $\textbf{t}_{1}$ (LSTs 0.5-4.6 hours) and $\textbf{t}_{2}$ (LSTs 
4.6-8.6 hours). We form our two datasets as $\textbf{x}_{1} = \textbf{t}_{1} + \textbf{t}_{2}$ and $\textbf{x}_{2} = \textbf{t}_{1} -\textbf{t}_{2}$.
\end{itemize}

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/null.png}
	\caption{Results for three null tests compared to the PAPER-64 unweighted revised power spectrum (black) and analytically estimated 
$2\sigma$ errors (gray shaded region). For each, we take a jackknife along a different axis of the dataset - along julian days (separating 
even and odd days; blue), along baselines (green), and along LST (magenta). We expect the sky signal to disappear for a `passing' null test (for all error bars to be contained to within the gray shaded region).
We find that none of the null tests performed pass, and that the dominant source of the biases in our power spectrum are likely caused by variation of foregrounds in LST (due to the magenta result being the most inconsistent with noise)}
	\label{fig:null}
\end{figure*}

Investigating Figure \ref{fig:null}, we find that none of the null tests are fully consistent with noise; we would expect the null tests' error bars to be contained within the gray shaded region (theoretical error) for a `passable' test. We note, however, that the error bars for the baselines test (green) and even/odd test (blue) are smaller than those of the LST test (magenta).

%We note that the error bars for the baselines test (green) are especially small --- this is because the signal is effectively cancelled out prior to cross-multiplication (when forming baseline groups).

This suggests that the primary systematic driving our power spectrum points and errors to be outside the 
estimated error region is LST-dependent, and therefore likely to be residual foregrounds. In other words, the fact that many magenta power spectrum points are outside the estimated error region implies that $t_{1}$ and 
$t_{2}$ differ by an amount greater than thermal noise and they do not effectively cancel out the sky signal during cross-multiplication. While residual foregrounds are the primary cause for power spectrum results above the theoretical noise (as in Figure \ref{fig:ps1_data}), other unexplained variations exist which are present in all the null tests. These variations could be attributable to... \cc{fill in}

>>>>

The highest sensitivity power spectrum result for PAPER-64 using the updated analysis presented in this paper, shown in 
Figure \ref{fig:ps1_data}, has positive biases at low $k$ values. As discussed in Section \ref{sec:BiasTypes}, these detections 
are most likely attributable to foreground leakage. The cause for the detections at higher $k$ values is more difficult to pinpoint. Here we demonstrate three null tests performed on PAPER-64 data that aim to explain these detections and verify 
that they are not attributable to EoR.

@@

\begin{itemize}
%\item Original Data (black): This is identical to the unweighted, revised PAPER-64 power spectrum in Figure \ref{fig:ps1_data} (one baseline type only) with weighting matrix $\textbf{I}$. There are clear detections at low $k$'s.
\item Even/Odd (blue): We split our dataset into even (\textbf{e}) and odd (\textbf{o}) days. We then form the following two datasets: $\textbf{x}_{1} = \textbf{e} + \textbf{o}$ and $\textbf{x}_{2} = 
\textbf{e} - \textbf{o}$. We form datasets in this way to ensure that we use the full sensitivity of our data. When cross-multiplied, we obtain:
\begin{equation}
\textbf{x}_{1}\textbf{x}_{2}^{\dagger} = \textbf{ee}^{\dagger} - \textbf{eo}^{\dagger} + \textbf{oe}^{\dagger} - \textbf{oo}^{\dagger}
\end{equation}
If the same sky signal is in both the even and odd datasets, we expect it to cancel out.
\item Baselines (green): We split our dataset into two halves, where each contains half of the total baselines used in the 
analysis. Again, we form $\textbf{x}_{1} = \textbf{b}_{1} + \textbf{b}_{2}$ and $\textbf{x}_{2} = \textbf{b}_{1} - \textbf{b}_{2}$, where $\textbf{b}_{1}$ is the first baseline 
group and $\textbf{b}_{2}$ is the second baseline group.
\item LST (magenta): We split our dataset into two halves based on LST, namely $\textbf{t}_{1}$ (LSTs 0.5-4.6 hours) and $\textbf{t}_{2}$ (LSTs 
4.6-8.6 hours). We form our two datasets as $\textbf{x}_{1} = \textbf{t}_{1} + \textbf{t}_{2}$ and $\textbf{x}_{2} = \textbf{t}_{1} -\textbf{t}_{2}$.
\end{itemize}

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/null_zoom.png}
	\caption{Results for three null tests compared to analytically estimated 
$2\sigma$ errors (gray shaded region). For each of the tests, we take a jackknife along a different axis of the dataset - along julian days (separating 
even and odd days; blue), along baselines (green), and along LST (magenta). We expect the sky signal to disappear for a `passing' null test.
We find that the baselines (green) and even/odd (blue) tests are consistent with noise, and that the dominant source of the biases in our power spectrum are likely caused by variation of foregrounds in LST (due to the magenta result being the most inconsistent with noise)}
	\label{fig:null}
\end{figure*}

Investigating Figure \ref{fig:null}, we find that the baselines test is fully consistent with noise, and that the even/odd test is mostly consistent with noise. For a `passing' null test, we expect $95\%$ of the error bars to be contained within the gray shaded region (theoretical error). We additionally note that the error bars for the baselines test (green) are especially small --- this is because the signal is effectively cancelled out prior to cross-multiplication (when forming baseline groups).

The null test in LST is much more illuminating. Because of the large error bars and variation in its result, this suggests that the primary systematic driving our power spectrum points and errors to be outside the 
estimated error region is LST-dependent, and therefore likely to be residual foregrounds. In other words, the fact that several magenta power spectrum points are outside the estimated error region implies that $t_{1}$ and 
$t_{2}$ differ by an amount greater than thermal noise and they do not effectively cancel out the sky signal during cross-multiplication.
